<style type="text/css">
.main-container {
  max-width: 1600px;
  margin-left: auto;
  margin-right: auto;
}
</style>


<style type="text/css">
body, td {
   font-size: 20px;
}
code.r{
  font-size: 16px;
}
pre {
  font-size: 16px
}
</style>

--- 
title: "San Francisco Trees"
author: "Omar Lopez Rubio"
subtitle: "Tidy Tuesday - 2020-01-28"
---

So this week we have some data on San Francisco trees.
This week I'm more interested in plotting some fancy maps than in anything else, so I researched a little bit how to achieve it. I found a package
called `leaflet`. It allows you to plot interactive maps using coordinates
and also add markers to it. The markers are fully customizable.
The only thing I didn't like is that I didn't find good docs about it.

```{r, message=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
knitr::opts_chunk$set(fig.width=12, fig.height=8) 

sf_trees <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-28/sf_trees.csv')

glimpse(sf_trees)

trees <- sf_trees %>%
  mutate_if(is.character, fct_explicit_na)

```

Basic data visualization and exploratory analysis

```{r}
trees %>%
  count(species) %>%
  arrange(n)
```

We have a lot of species with less than 100 observations, may be worth to 
discard those samples.
Let's take a look at `legal_status`

```{r}
trees %>%
  count(legal_status, sort = TRUE)
```
Here we can see that it happens also with legal_status.
I'm going to do some `geom_col()` plots for them and also the other variables, to get a better idea of the dataset.

```{r}

trees %>%
  count(species, sort = TRUE) %>%
  head(20) %>%
  mutate(species = reorder(species, n)) %>%
  ggplot(aes(x = species, y = n)) + geom_col() + coord_flip()


trees %>%
  count(legal_status, sort = TRUE) %>%
  mutate(legal_status = reorder(legal_status, n)) %>%
  ggplot(aes(x = legal_status, y = n)) + geom_col() + coord_flip()

trees %>%
  count(site_info, sort = TRUE) %>%
  mutate(site_info = reorder(site_info, n)) %>%
  ggplot(aes(x = site_info, y = n)) + geom_col() + coord_flip()

trees %>%
  count(caretaker, sort = TRUE) %>%
  mutate(caretaker = reorder(caretaker, n)) %>%
  ggplot(aes(x = caretaker, y = n)) + geom_col() + coord_flip()
```

So visually we can see that there are some factors that it's count is really low.
For demonstration purposes, I'm going to use `fct_lump` to collapse a little bit the species column.

```{r}
trees_lump <- trees %>% 
  mutate(species = fct_lump_min(species, 100))

trees_lump %>% 
  count(species, sort = TRUE)

trees_lump %>%
  filter(species == "Other")

```

I'm interested in check wheter the caretaker is private or not, so I'm going to create a new column to check it, and later make a model with `tidymodels`.

```{r}
trees_proc <- trees_lump %>%
  mutate(isPrivate = case_when(
    as.character(caretaker) == "Private" ~ as.character(caretaker),
            TRUE ~ "Other")) %>%
  filter(longitude > -125) %>%
  select(-caretaker)

trees_proc %>%
  count(isPrivate, sort = TRUE)
```

Cool! Let's end the visualization making a plot with the `lat` and `long`
For the sake of complexity and visualization, I'm only going to plot 2000 samples in the interactive plot, and 10000 in the static one.

```{r}
library(leaflet)

set.seed(555)
pal <- colorFactor(c("green", "navy"), domain = c("Other", "Private"))
trees_proc %>%
  sample_n(2000) %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(
    radius = 0.5,
    color = ~pal(isPrivate)
  ) %>%
  addLegend("bottomright", pal = pal, values = ~isPrivate,
             title = "Caretaker type")

trees_proc %>%
  sample_n(10000) %>%
  ggplot(aes(x=latitude, y = longitude, color = isPrivate)) + geom_point(alpha = 0.2) + theme_minimal()

```
Time to model things!
I'm going to downsample `isPrivate`, because we are working with an unbalanced dataset.
Also I'm gonna do a little bit more of preprocessing,
`recipes` makes it really easy.
In Julia Silge's screencast I saw how she used the `doParallel` library, 
and after checking the package out I can't figure why I didn't use it until now.
The training may take a while...
```{r, message=FALSE}
library(tidymodels)
library(lubridate)

trees_proc <- trees_proc %>%
  mutate(site_order = as_factor(site_order)) %>%
  mutate(site_order = fct_explicit_na(site_order, na_level = "-1"))

trees_split <- initial_split(trees_proc)
trees_train <- training(trees_split)
trees_test <- testing(trees_split)

trees_recipe <- trees_train %>%
  recipe(isPrivate ~ .) %>%
  update_role(address, new_role = "loc") %>%
  update_role(tree_id, new_role = "id") %>%
  step_rm(date, plot_size, dbh) %>%
  step_downsample(isPrivate) %>%
  step_dummy(all_predictors(), -all_outcomes(), -all_numeric()) %>%
  prep()
  

trees_train <- juice(trees_recipe)
trees_testing <- trees_recipe %>%
  bake(trees_test)

library(doParallel)
doParallel::registerDoParallel()
trees_ranger <- rand_forest(trees = 100, mode = "classification") %>%
  set_engine("ranger", importance = "permutation") %>%
  fit(isPrivate ~ ., data = trees_train)

predict(trees_ranger, trees_testing)

trees_ranger %>%
  predict(trees_testing) %>%
  bind_cols(trees_testing) %>%
  metrics(truth = isPrivate, estimate = .pred_class)

trees_ranger %>%
  predict(trees_testing) %>%
  count(.pred_class)
```


